\section{Existing \cl{} readers}

In this section, we give an overview of the native readers of some
common existing \cl{} implementations.  We analyze each one with
respect to its behavior in error situations, and with respect to
features that can be essential in various use cases, including cases
where the reader is not used as the first phase of a
\texttt{read-eval-print} loop.

\subsection{Embeddable Common Lisp (\ecl{})}

\ecl{} is implemented as a core written in \clanguage{}.  It is built by
first compiling this core and then adding code written in \cl{} by
loading it into the system resulting from the compilation of the core.
For that reason, the reader must be written largely in \clanguage{}.

When an attempt is made to read an incomplete expression, so that the
\texttt{end-of-file} condition is signaled, no restart other than
\texttt{restart-toplevel} is available.

Similarly when an attempt is made to read a symbol with an explicit
package marker where the corresponding package does not exist, an
error of type \texttt{simple-error} is signaled, and again, the only
restart available is \texttt{restart-toplevel}.

When an attempt is made to read an incorrect token in the form of a
malformed symbol with three package markers, again an error of type
\texttt{simple-error} is signaled, and the message is cryptic.  For
example, an attempt to read the malformed symbol \texttt{cl:::car},
the error message is ``There is no package with the name
\texttt{CL:::C}''.

One might think that the reason \ecl{} proposes so few restarts is
because it might difficult to do so for a function written in
\clanguage{} but that is not the case; \ecl{} has full support for
functionality analogous to \texttt{restart-case} for \clanguage{}
code.

The \ecl{} reader is not involved in source tracking.  The compiler
tracks the source only of top-level forms, so recursive calls to the
reader need not be handled.

\subsection{Steel-Bank Common Lisp (\sbcl{})}

The \sbcl{} reader is written in \cl{}, without the use of the specific
features of the \cl{} Object System (\clos{}) such as generic
functions and standard classes.

When an attempt is made to read an incomplete expression, so that the
\texttt{end-of-file} condition is signaled, no restart other than
\texttt{retry} and \texttt{abort} is available.

When an attempt is made to read a symbol with an explicit package
marker where the corresponding package does not exist, an error of
type \texttt{simple-reader-package-error} is signaled.  Several
restarts are then available:

\begin{itemize}
\item using the current package instead,
\item retry finding the package (presumably after it has been created
  by the user),
\item specify a different package,
\item return an uninterned symbol, and
\item specify a symbol to return.
\end{itemize}

\noindent
When an attempt is made to read an incorrect token in the form of a
malformed symbol with three package markers, an error of type
\texttt{simple-reader-error} is signaled, and no restart other than
\texttt{retry} and \texttt{abort} is available.

The \sbcl{} reader provides several extensions.  One such extension is
that a package prefix with two colon characters can precede any
expression, and not just a symbol.  The effect is that the reader then
uses the package named by the prefix as the default package to intern
symbols in the expression.

Another extension is that the exponent marker \texttt{r} or \texttt{R}
can be used to return a rational number that is the exact value of the
floating-point syntax used.

The names of symbols created by the \sbcl{} reader can be normalized
according the ``Normalization Form KC'' which is specified for
Unicode.  Whether symbol names are normalized in this way is
controlled by the accessor \texttt{readtable-normalization} which
operators on a readtable.

The \sbcl{} reader can track source positions of each object returned.
Source tracking is handled through a special type of stream.  The
function \texttt{\%read-preserving-whitespace} checks for the stream
argument being of this type and, if so, registers the form and the
position of the form.  Two parallel adjustable vectors are used for
this purpose.

\subsection{Clozure Common Lisp (\ccl{})}

The \ccl{} reader is written in \cl{}, without the use of the specific
features of the \cl{} Object System (\clos{}) such as generic
functions and standard classes.

When an attempt is made to read an incomplete expression, so that the
\texttt{end-of-file} condition is signaled, only restarts that return
to previous break levels are available.

When an attempt is made to read a symbol with an explicit package
marker where the corresponding package does not exist, several
additional restarts are available:

\begin{itemize}
\item supplying a different name for the package
\item retry finding the package (presumably after it has been created
  by the user),
\item making the supplied package name a nickname for the package
  \texttt{common-lisp-user}.
\end{itemize}

\noindent
When an attempt is made to read an incorrect token in the form of a
malformed symbol with three package markers, no error is signaled at
all, and instead the third package marker is treated as escaped, so
that it is part of the symbol.

The \ccl{} reader can track source positions.  It registers the stream
position before and after an expression has been read, and creates an
instance of a \texttt{struct} named \texttt{source-note} and stores
this instance in a hash table.  Source positions are recorded only if
the stream is an element of the list
\texttt{*recording-source-streams*}, which is the case when the reader
is called for the purpose of evaluating the resulting expression.

\subsection{Clasp}

Clasp is a \cl{} implementation that was specifically designed to
interoperate with \cplusplus{} code, in order to take advantage of
existing libraries written in \cplusplus{}.  Clasp was originally
written as a conversion of the \clanguage{} code of \ecl{} to
\cplusplus{} and with almost all the \cl{} code coming from \ecl{}.
As such, its reader is a \cplusplus{} conversion of the reader from
\ecl{}.  However, in later bootstrapping stages, Clasp uses Eclector,
the subject of this paper, in order to make Eclector the default
reader.

There are plans to modify the bootstrapping procedure of Clasp so that
the initial \cplusplus{} reader is no longer needed.  For all these
reasons, we do not review the \cplusplus{} reader of Clasp, and
Eclector is fully described in \refSec{sec-eclector}.

\subsection{LispWorks}

LispWorks is a commercial \cl{} implementation, so we obviously do not
have access to its source code.  For that reason, we cannot determine
how the reader is written.  We can, however, investigate what restarts
are available in various situations.

When an attempt is made to read an incomplete expression, so that the
\texttt{end-of-file} condition is signaled, only the \texttt{abort}
restart is available.

When an attempt is made to read a symbol with an explicit package
marker where the corresponding package does not exist, several
additional restarts are available:

\begin{itemize}
\item retry finding the package (presumably after it has been created
  by the user),
\item use another package,
\item create the package, and
\item FIXME: and what?  Did I mean to put something here?
\end{itemize}

\noindent
When an attempt is made to read an incorrect token in the form of a
malformed symbol with three package markers, no restart other than
\texttt{abort} is available.

The LispWorks manual does not contain an entry for \texttt{read},
which we take to mean that the user cannot configure the reader in any
way other than what the \cl{} standard allows.

\subsection{Scheme}

Many implementations of the \scheme{} language support the concept of
\emph{syntax objects} which are \scheme{} datums annotated which
additional information such as scope information for identifiers or
source information for forms read from files.  While the original
motivation for syntax objects seems to have been hygienic macros,
attaching source information to all datums\footnote{not just lists,
  vectors, etc.} is clearly useful as well.


See for instance \texttt{read-syntax} in the Guile manual.%
\footnote{https://www.gnu.org/software/guile/manual/html\_node/Annotated-Scheme-Read.html}

As far as we can tell, syntax objects allow and actually are used for
reading \emph{forms} with source information but don't seem to support
producing parse results for \emph{skipped material} like comments.

Possible Limitation: When trying to read something circular in Racket it says
\texttt{read-syntax: `\#...\#` forms not enabled for `read-syntax`}
