\section{Conclusions}

\subsection{Challenges}

As remarked at the end of \refSec{sec:benchmarks}, when executed in a
given \cl{} implementation, \eclector{} tends to run 4 (\sbcl{}) to 10
(\ccl{}) times slower than the respective native reader which is a
significant slowdown.  We suspect that two major factors cause the
difference in performance:
\begin{enumerate}
\item \eclector{}'s portable code cannot use implementation specific
  optimizations that are available to native readers.
\item The numerous possible ways in which \eclector{} can be
  configured via generic functions slow down the execution, in
  particular when the native reader does not use generic functions
  because ordinary functions perform better in the implementation in
  question.
\end{enumerate}
We believe that \eclector{}'s performance characteristics are still
acceptable for most use cases: When a reader is used as part of file
compilation or a \texttt{read-eval-print} loop, the time taken by the
reader is usually dwarfed by the time taken by to compiler or
evaluator.  Conversely, for interactive uses like parsing an editor
buffer at typing speed, no reader with subsequent analysis steps will
be fast enough to process an entire buffer, with arbitrarily sized
contents, quickly enough.  In such a situation, \eclector{} can be
used \emph{incrementally} as described in
\refSec{sec:incremental-parsing} so that the amount of parsing work
after each change is approximately constant.

% TODO: explain this: raw performance may be not as good as other
% readers but we try to avoid problems that stem from algorithms with
% bad asymptotic complexity so that there are no pathological inputs.

Another challenging area is forward and backward compatibility,
particularly in two areas: customization possibilities provided to
clients and reader macros.  For clients, in particular since
\eclector{} already has some users, it is important to not introduce
breaking changes in customization protocols.  On the other hand, when
\eclector{}'s customization possibilities prove insufficient for some
new use case, the protocols have to be changed or extended.  Reader
macros, which are typically not aware of \eclector{} and its
facilities, can be at odds with many of the goals mentioned above such
as avoiding effects, always being able to recover, etc.  As a
secondary issue, more customizability and more sophisticated
mechanisms lead to a widening gap between reader macros that make use
of \eclector{}'s facilities and fully portable reader macros.

% User-written (S-expression) macros can be at odds with source
% tracking.  The concrete syntax tree library provides
% \texttt{reconstruct} but that is approximate.  Other approaches
% include CST-based or AST-based macros and Scheme-like syntax objects
% but those approaches would be very disruptive for \cl{} users
% (programmers).

% Slow development speed

\subsection{Future work}

We are working on the following new customization protocols for
\eclector{}:

The \cl{} reader associates \emph{constituent traits} with characters
that occur within symbols and numbers.  The traits associated with
each character are fixed and listed in the \cl{} specification.  For
\eclector, we plan to implement a ``traits table'' that allows clients
to change the constituent traits associated with each character.

We are working on a customizable and optimizing quasiquotation
processor for \eclector{}.  Customizable here means, for example, that
quasiquotation support can be added for new literals such as
hash-table literals.  Optimizing means that quasiquoted material will
be turned into forms that can be evaluated efficiently.

We plan to generalize the protocol for literal construction which
currently consists of a few scattered functions like
\mintinline{lisp}{find-character},
\mintinline{lisp}{make-structure-instance} to a more general signature
like \mintinline{lisp}{make-literal} that would also handle numbers,
symbols, arrays and other literals.  This change will also incorporate
the Quaviver
library\footnote{\url{https://github.com/s-expressionists/Quaviver/}}
for building floating point numbers as well as better handling of
float overflows.

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: "eclector"
%%% End:
