\section{Eclector: A highly configurable reader}
\label{sec-eclector}

In this section, we discuss features and other characteristics of
\eclector{}.  \eclector{} was initially created as the native reader
for the \sicl{}%
\footnote{https://github.com/robert-strandh/SICL}
project, still under development.  From the start, it was designed to
make it possible to parse an editor buffer as described in
\cite{durand:hal-01887230}.  Early on, the code was extracted and
renamed to \eclector{}.  From then on, it evolved entirely
independently of \sicl{}, through the work of the first author of this
paper. 

\subsection{Features}

\subsubsection{Source information}

\eclector{} can be called in a way that it produces a \emph{concrete
syntax tree} which represents the S-expression so that each node is
wrapped in an instance of a standard class.  In addition to the raw
S-expression, this instance contains source information, i.e., the
position of the beginning and the end of the wrapped expression in the
input stream.  By default, the source information consists of the
stream position, but client code can provide a method on the relevant
generic function so that some other representation is used.  For
example, in \sicl{}, the source information consists of the entire
input file in the form of a vector of lines, and the line and column
of the start and the end of the S-expression.

\subsubsection{Skipped input}

\eclector{} can produce \emph{parse results} that contain information
that is normally skipped by a reader, such as comments, reader
conditionals, and invisible features in the input, such as labeled
object definitions and references, and feature expressions in reader
conditionals.

While skipped material serves no purpose when a reader is used as the
first step of a \texttt{read-eval-print} loop, it is essential when
the reader is used to parse the contents of an editor buffer.

\subsubsection{Specific error condition types}

For each error situation, \eclector{} signals a condition of a type
that is specific to that situation.  That way, client code can handle
each such situation differently.

\subsubsection{Recovery from errors}

For every error situation, restarts are available that allow client
code to indicate to \eclector{} that it should recover from the error.

If an \texttt{end-of-file} condition is signaled, recovery consists of
acting as if input is complete, so that string literals, lists,
vectors, etc., are closed.

If a the name of a non-existing package is used as a package prefix
for a symbol, a restart can be invoked that returns an uninterned
symbol instead.

\subsubsection{Intrinsic and extrinsic use}

\eclector{} can be loaded into an existing \cl{} implementation
without affecting the native reader, or any other system-specific
code.  It will then define several \eclector{}-specific packages that
a client must refer to in order to use \eclector{} this way.  We call
this way of using \eclector{} ``extrinsic''.

But \eclector{} can also be used as the native reader of a new \cl{}
implementation, or to replace the existing reader of an existing \cl{}
implementation.  Its main functions will then be defined in the
\texttt{common-lisp} package.  We call this way of using \eclector{}
``intrinsic''.  Intrinsic use in a new \cl{} implementation will
require some specific bootstrapping consideration, since
\clos{} is used extensively both in the implementation of \eclector{}
for configuring it.

\subsubsection{Reader state protocol}
\label{sec:reader-state-protocol}

Any native \cl{} reader (and the intrinsic version of \eclector{} does
this as well) consults several special variables in order to determine
actions in various situations.  These special variables are set by
user code, by other features of the \cl{} system, and (rarely) by the
reader itself.  For example, the reader uses the special variable
\texttt{*readtable*} to determine which readtable to use to determine
the \emph{syntax type} associated with a character.  Another example
of such a special variable is \texttt{*read-base*} that determines the
numeric base for reading numbers.

When \eclector{} is used extrinsically, it is clearly not possible to
use the host versions of these variables.  For that reason, among
others, \eclector{} defines a \emph{reader state protocol} that
replaces the direct use of these special variables and.  Instead, the
reader state protocol uses \emph{symbols} (often the same symbols that
name the host special variables) called \emph{aspects} to denote the
variables.

The reader state which \eclector{} queries and manipulates through the
reader state protocol contains aspects that do not denote any of the
standard special variables: one aspect controls whether the consing
dot is legal in the current state.  Another aspect tracks the
quasiquotation nesting level.  User code can in principle add
additional, custom aspects.  The intention of the reader state
protocol is to allow representing the entire state of the reader as
one or more first-class objects (TODO: talk more about how that helps
with incremental parsing?).

The reader state protocol consists of a small number of generic
functions like \texttt{state-value}, \texttt{(setf state-value)} and
\texttt{call-with-state-value} that are called by \eclector{} and
client code to query or modify the aspects of the reader state.  Each
of these generic functions will have an aspect as one of its
arguments.  These functions also have a \emph{client} parameter that
can be used by client code to create specific methods.

We could talk about the following unusual features:

\begin{itemize}

\item Most aspect of the reader are customizable. Customization in
  standard \cl{}: Reader control variables and Readtable (case, syntax
  type, macros)

\end{itemize}

\subsection{Robustness and Unusual Inputs}

\eclector{} uses a ``hybrid recursion'' approach in order to be able
to process deeply nested expressions containing reader labels and
references to those labels as well as parse results of similar
structure.  The approach consists of initially processing the input
recursively but tracking the recursion depth and employing iteration
and a worklist in case the recursion depth exceeds a threshold.
Typical readers (like that in \sbcl{} for example) may fail, with a
stack overflow, to read and process some such expressions, whereas
\eclector{} succeeds.  See \refApp{app-nesting} for an example of such
an expression.

Not all recursion in \eclector{} is handled using the ``hybrid
recursion'' approach at the moment.  In particular, \texttt{read}
calls cannot nest to an arbitrary depth with the current design.  It
may be possible to leverage the reader state protocol
\seesec{sec:reader-state-protocol} to break up these nested calls into
chunks manageable by recursive processing but we have not attempted
this technique.

We can talk about the ``hybrid recursion'' strategy for handling
deeply nested structures and similar issues.  We could give examples
that, say, \sbcl{} cannot handle.

\subsection{Efficiency}

Clearly, the numerous possible ways in which \eclector{} can be
configured affects performance.  And on \cl{} implementations with
poor performance for generic dispatch, performance is going to be
affected even more, since \eclector{} configuration is handled mostly
by methods on generic functions.

However, when a reader is used as part of a \texttt{read-eval-print}
loop, the performance of the reader has very little importance, since
the time it takes to read a form is dwarfed by the time it takes to
compile or evaluate the result.  There are of course other situations
where the reader is used as such, but then the native reader is
probably unnecessarily configurable as well in many situations.

We also use \eclector{} to parse the contents of an editor buffer at
typing speed, and it would seem that performance is very important in
this situation.  However, we then use the reader \emph{incrementally}
and described the paper presented at ELS'18
\cite{durand:hal-01887230}.  Furthermore, it would be impossible to
use a native reader for this use case since many of the configurations
that \eclector{} allows are absolutely necessary.

We could do a few benchmarks

Another thing we could talk about are edge cases of labeled objects
and references.  Not because the performance of those is particularly
relevant in practice but because there is algorithmic innovation in
\eclector{} in this area and we even have an edge over \sbcl{} for
certain cases.

% LocalWords:  uninterned

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: "eclector"
%%% End:
