\section{Applications}
\label{sec-applications}

In this section, we describe use cases that benefit from features of
\eclector{} like extensibility, source position tracking and error
recovery.  We also list applications of \eclector{} in existing \cl{}
projects that are concrete examples of these use cases.

\subsection{Use cases}

\subsubsection{Intrinsic use: using \eclector{} as the reader of a \cl{} implementation}

The most obvious application for \eclector{} is as a native reader of
some \cl{} implementation.  Currently, \eclector{} is used this way
only in Clasp, but it will also be used in the \sicl{} \cl{}
implementation, once it is finished.

Currently, \sicl{} uses \eclector{} extrinsically during
bootstrapping, in order to read forms to be evaluated in a first-class
global environment \cite{Strandh:2015:ELS:Environments}.  Since the
host global environment must not be affected as a result of reading
these forms, we take full advantage of the ability to configure
\eclector{} to avoid such side effects.

\subsubsection{Alternate syntax beyond readtable and reader macros}

https://github.com/pve1/eclector-access/

Coalton language: https://github.com/coalton-lang/coalton/blob/main/src/parser/reader.lisp

maybe Hayley's language

\subsubsection{Reader part of sandboxed evaluation}

The goal of \emph{sandboxed evaluation} is to (read and) evaluate untrusted code, for
example code received from possibly adversarial third parties, in way
that makes it impossible to affect the surrounding environment
except in limited, explicitly permitted ways.  examples: errors leading to denial of service, read-time eval, interning,
consuming resources such as stack space, heap space, processing

\subsubsection{Static Analysis}

For some applications doing \emph{static analysis} of \cl{} code, it
is not possible, or at least undesirable, to use the native reader.  A
typical example would be a so-called \emph{linter}, i.e., a tool for
analyzing code with the purpose of giving the user more feedback about
the code than a typical compiler will do.  Such feedback can include
violations of conventional style and suggestions for improving the
readability of the code.  Other examples include

\begin{itemize}
\item spelling mistakes in comments or symbol names
\item unnecessary package prefix in case the symbol in question is
  accessible in the current package or a package-local nickname allows
  a shorter spelling.
\item use of two-package markers when spelling an external symbol
\item use of the wrong number of semicolons for a comment at a given
  level (for example two semicolons for a comment at the toplevel)
\item logical errors in reader conditionals such as
  \texttt{\#+win32(\dots{} \#+(and thread (not win32))(\dots{}) \dots{})}
\end{itemize}

Such tools should preferably not have any side effects on the host
\cl{} image such as:

\begin{itemize}
\item interning symbols in some host package,
\item altering the package system, or
\item evaluating code.
\end{itemize}

The host reader can then not be used because of the possibility of
read-time evaluation, or simply the use of double package markers for
some symbol that will then be interned in some host package.  And the
code being analyzed may contain package markers for packages that have
not been created in the host \cl{} system.

An important feature of a reader to be used with such a tool is the
desire to recover from errors such as syntax errors in tokens.  The
tool would then report the error and then continue the analysis.

In some cases ``invisible'' objects such as feature conditionals,
labeled objects and skipped material like comments when those things
are the subject of the analysis.

Important features for static analysis

\begin{itemize}
\item No side effects (Interning, package system changes, code evaluation)
\item Error recovery
\end{itemize}

Limitations

\begin{itemize}
\item Reader macros
\item Read-time evaluation
\end{itemize}

\subsubsection{Syntax highlighting}

\todo{Screenshot}

For documentation

Important features

\begin{itemize}
\item No side effects (Interning, package system changes, code evaluation)
\item Parse results, skipped material, ``invisible'' objects
\item Error recovery
\end{itemize}

\subsubsection{Incremental parsing for editor-like applications}
\label{sec:incremental-parsing}

\eclector{}'s ability to recover from errors, represent the state of
the reader as a first class object, produce parse results for
expressions as well as skipped material with source information and
operate without side-effects enable incremental parsing.  Incremental
parsing is useful in interactive, editor-like applications like REPLs,
Listeners, IDEs, LSP\footnote{Language Server Protocol}-servers and of
course source code editors.

\begin{itemize}
\item No side effects (Interning, package system changes, code evaluation)
\item Parse results, skipped material, ``invisible'' objects
\item Reader state protocol
\item Error recovery
\end{itemize}


\subsection{Projects using \eclector{}}

\subsubsection{formgrep}

The library \texttt{formgrep}%
\footnote{\url{https://github.com/death/formgrep}}
provides functionalities similar to those of the \unix{} \texttt{grep}
utility, except that it works on S-expressions rather than text.  This
library takes a collection of files and a regular expression as
arguments.  It then uses \eclector{} to parse each file.  It defines a
method on the \eclector{} generic function \texttt{interpret-symbol}
that uses the regular expression on the token representing a symbol
that \eclector{} recognizes to determine whether there is a match.  If
so, it accumulates the token in a list of matches that is ultimately
returned.  For reader conditionals, \texttt{formgrep} configures
\eclector{} to always read the guarded expression regardless of the
condition so that no part of the code is skipped and thus excluded
from the search.  \texttt{formgrep} uses the idiom
\mintinline[fontsize=\small,escapeinside=||]{lisp}{(handler-bind ((error |\#|'eclector.base:recover)) |\dots| (read |\dots|) |\dots|)}
to deal with parts of the input that are either invalid or cannot be
read with the employed \eclector{} configuration.

% \subsubsection{rewrite-cl}

% The library \texttt{cl-rewrite}%
% \footnote{https://github.com/atgreen/rewrite-cl}
% was written by Anthony Green.  It
% parses \cl{} code as text into a structure called a \emph{zipper}.
% Each node in the zipper contains a corresponding element of the
% S-expression being parsed.  The type of the node corresponds to the
% nature of the element, such as a token, a list, a vector, etc.
% Furthermore, each node contains information about the source position
% of the element.

% The zipper structure can then be manipulated in various ways, such as
% nodes being replaced, and the result can then be written to a file.

\subsubsection{mallet}

The purpose of the library \texttt{mallet}%
\footnote{\url{https://github.com/fukamachi/mallet}} which was written
by Eitaro Fukamachi is to read \cl{} code and analyze it according to
configurable rules in order to detect mistakes and deviations from a
desired coding style.  The library configures \eclector{} so that it
behaves differently compared to the standard reader in multiple ways:

\begin{enumerate}
\item \eclector{} is configured to produce parse results with source
  information.
\item \texttt{mallet} recovers from many errors by specifically
  handling a number of the specialized conditions which \eclector{}
  signals.
\item \texttt{mallet} handles elements of the code that cannot be read
  in the usual way without affecting the Lisp image such as symbols
  that would require interning or read-time evaluation by configuring
  \eclector{} to produce placeholder objects instead.
\end{enumerate}

\subsubsection{Second Climacs editor}

\todo{capitalization of section}
As mentioned in \refSec{sec-eclector}, \eclector{} was designed from
the start to be able to parse the contents of an editor buffer.  We
use the ability to do incremental parsing in the Second Climacs text
editor, still under development.  The objective is to have an editor
specifically for writing \cl{} code that can analyze the buffer
contents more faithfully on the level of character syntax,
s-expression syntax and semantics than existing editors can do.  The
character syntax is handled by a combination of \eclector{} and the
incrementalist library%
\footnote{\url{http://github.com/s-expressionists/incrementalist}}
which uses an extension of the incremental parsing described in
\refSec{sec:incremental-parsing}.  The s-expression syntax is handled
by a separate library that accepts the concrete syntax trees produced
by \eclector{} and applies grammar rules which describe the \cl{}
operators.

\subsubsection{Other projects which use \eclector{}}

The following projects also use \eclector{} in some way:

\begin{itemize}
\item \texttt{common-lisp-jupyter}\footnote{\url{https://github.com/yitzchak/common-lisp-jupyter}}
\item \texttt{cl-mcp}\footnote{\url{https://github.com/cl-ai-project/cl-mcp}}
\item \texttt{Software Evolution Library}\footnote{\url{https://github.com/GrammaTech/sel}}
\item \texttt{40ants-critic}\footnote{\url{https://github.com/40ants/40ants-critic}}
\item \texttt{proto}\footnote{\url{https://github.com/BnMcGn/proto}}
\item \texttt{yew}\footnote{\url{https://github.com/lispnik/yew}}
\item \texttt{is-dead-fossil}\footnote{\url{https://github.com/adlai/is-dead-fossil}}
\item \texttt{staple}\footnote{\url{https://codeberg.org/shinmera/staple/src/branch/master/parser/walker.lisp}}
\item \texttt{apprentice}\footnote{\url{https://github.com/pve1/apprentice}}
\end{itemize}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: "eclector"
%%% End:
